---
title: "Bad weather Kelheim Demo"
author: "Oleksandr Soboliev"
#always_allow_html: true
output:
  #pdf_document: default
  html_notebook:
    theme: cosmo
    highlight: monochrome
    code_folding: show
editor_options: 
  chunk_output_type: inline
---
```{r, include= FALSE}
library(tidyverse)
library(lubridate)
library(plotly)
```


## **Research Meteostat**
After some researches about meteostat data nearest station in DE that belongs to Kelheim region 
are:

* "Mallersdorf-Pfaffenberg/Niederbayern" with an id: "D3147"
* "Neumarkt / Höhenberg" with and id: "69110"
* Uebungsdorf / Emhof

there are many of them so I am starting to think about extracting all from the Bayern or extract the nearest from longtitude/latitude point with the Kelheim shapefile(using json and Euclid distances) 

[Kelheim has no weather station, but it could be reconstructed with 2 other](https://weatherspark.com/y/70370/Average-Weather-in-Kelheim-Germany-Year-Round)

Hohenfels with id: "10775" and Ingolstadt with id:"10860"
**kelheim_data = {weight1}x{hohenfels} + {weight2}x{inglstadt}**

Also this site shows, that there are many of the Kelheim stations in this area, but meteostat doesn't contain them
https://www.wunderground.com/dashboard/pws/IKELHE5



## **Research Weatherstack**

```{r first look at weatherstack data specific to Kelheim}
weatherstack_kelheim = read_delim("data/Kelheim_weather_since_july_2008.csv",delim = ",")
print(weatherstack_kelheim)
```
What to take as a reffer point isn't clear because of the date(before/after covid) and weather type (sunny,clear,temperature)
Also there is no temperature in it :/

## **Import mobility from Google**
```{r including google germany mobility data,message=FALSE}
global_mobility = read_delim("https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv",",")
de_mobility = global_mobility %>% filter(country_region_code == "DE")
```

```{r what regions are data provided}
print(unique(de_mobility$sub_region_1))

```
As we can see the most precise region to filter data from is Bavaria :/

Relevant data for the , mobility

```{r mobility data bavaria}
bavaria_mobility = de_mobility %>% filter(sub_region_1 == "Bavaria")
bavaria_mobility = bavaria_mobility %>% select(country_region,sub_region_1,date,residential_percent_change_from_baseline) %>%
  mutate(residential_percent_change_from_baseline = -residential_percent_change_from_baseline,
         source = "Google")%>%
  rename(BundeslandID = sub_region_1,not_at_home_change = residential_percent_change_from_baseline)
bavaria_mobility = bavaria_mobility %>% select(date,BundeslandID,not_at_home_change,source)
#Need to filter out weekends

plt = ggplot(bavaria_mobility)+
  geom_point(aes(x = date,y = not_at_home_change))
ggplotly(plt)
```
## **Import mobility from Senozon**

```{r import from senozon}
snz_mobility = read_delim("data/mobilityData_OverviewBL_weekly.csv",";")
snz_mobility = snz_mobility %>% filter(BundeslandID == "Bayern") %>% mutate(source = "senozon") %>% select(-outOfHomeDuration) %>% rename(not_at_home_change = percentageChangeComparedToBeforeCorona)
snz_mobility$date = as.Date(strptime(snz_mobility$date,"%Y%m%d"))
plt = ggplot(snz_mobility)+
  geom_point(aes(x = date,y = not_at_home_change))
ggplotly(plt)
```



## **Aggregate 2 sources**

```{r google+senozon}
mob_joined = rbind(snz_mobility,bavaria_mobility)
weatherstack_kelheim_precip_by_day = weatherstack_kelheim %>% group_by(date) %>% summarize(precip_mean = mean(precip))
mob_joined_with_weater = mob_joined %>% inner_join(weatherstack_kelheim_precip_by_day, by = "date")

plt = ggplot(mob_joined_with_weater)+
  geom_point(aes(x = date,y = not_at_home_change,color = source,size = precip_mean))
  

ggplotly(plt)

```


```{r shiny for html report}

```