---
title: "Bad weather Kelheim Demo"
author: "Oleksandr Soboliev"
output:
  html_document:
    df_print: paged
    code_folding: "hide"
runtime: shiny
editor_options:
  chunk_output_type: inline
---

```{r, include= FALSE}
library(tidyverse)
library(lubridate)
library(plotly)
library(leaflet)
library(rmarkdown)
library(modelr)
library(splines)
library(forecast)
library(fitdistrplus)
library(rjson)
```

# Data

Data is taken from these resources:

* Ingolstadt Weather from Meteostat [Meteostat](https://bulk.meteostat.net/v2/) 
* Weather Description from weatherstack [Weatherstck](https://svn.vsp.tu-berlin.de/repos/shared-svn/projects/KelRide/data/badWeather/weatherstack/)
* Mobility data is represented by number of Kexi Rides in Landkreis Kelheim (/shared-svn/projects/KelRide/data/KEXI/2021-04)
* Stringency (strictness of covid policies) number is from [Oxford COVID-19 Government Response Tracker](https://covidtracker.bsg.ox.ac.uk/)
* Holiday days are from [Feiertage](https://feiertage-api.de/)

# Regression analysis resources

Analysis was proceeded using this statistical sources:

* Linear Models with R (Julian J. Faraway)

# Importing and preparing data

Main target is to get result table containing all "expected" relevant data/variables that can be connected with changes of demand in Kelheim region, thus mobility data is collected on daily basis, remaining data has to be daily based.

After obtaining such table it would be useful(in sense of building a model and understanding the data) to check correlations between individual independent variable to dependent mobility variable. So the next chapter has a focus on filtering out unrelevant data to pass it to the model.

```{r importing all the data, message=FALSE,echo=FALSE,warning=FALSE}
# Ingolstadt weather
ingolstadt_weather = read_delim("https://bulk.meteostat.net/v2/daily/10860.csv.gz",",",col_names = FALSE)
colnames(ingolstadt_weather) = c("date", "tavg", "tmin", "tmax", "prcp", "snow", "wdir", "wspd", "wpgt", "pres", "tsun")

# Weatherstack data
weatherstack_kelheim = read_delim("data/Kelheim_weather_since_july_2008.csv",delim = ",")

# Stringency
json = fromJSON(file = "data/2022-10-08.json")
json = unlist(json)

#Mobility
demand = read_delim("data/allDemandByDate.csv")

#Holidays
holidays2020 = read_csv2("data/Holidays2020.csv") %>% dplyr::select(1,2,3)
holidays2021 = read_csv2("data/Holidays2021.csv") %>% dplyr::select(1,2,3)
holidays2022 = read_csv2("data/Holidays2022.csv") %>% dplyr::select(1,2,3)
holidays = rbind(holidays2020,holidays2021,holidays2022)
holidays = holidays %>% mutate(EndDateTime1 = as.Date(mdy_hm(EndDateTime1)),
                               StartDateTime1 = as.Date(mdy_hm(StartDateTime1)))

holiday_days = c(seq(holidays$StartDateTime1[1],holidays$EndDateTime1[1],by = "days"))

for(i in 1:nrow(holidays)){
  holiday_days = append(holiday_days,seq(holidays$StartDateTime1[i],holidays$EndDateTime1[i],by = "days"))
}

df_holidays = data.frame(date = holiday_days,isHoliday = TRUE)

```
##Modify and join data

```{r modify, message=FALSE, warning=FALSE}
# Weatherstack
weatherstack_kelheim_daily = weatherstack_kelheim %>%
  group_by(date) %>%
  count(description)

# Stringency 
deu_stringency = json[grep("DEU.stringency_actual",names(json))]
date_stringency = sapply(strsplit(names(deu_stringency),split = ".",fixed = TRUE),"[[",2)
df_stringency = data.frame(date = date_stringency,stringency = deu_stringency)
df_stringency = df_stringency %>% mutate(stringency = as.numeric(stringency), date = as.Date(date))



# Ingolstadt
type_of_weather = unique(weatherstack_kelheim$description)
map_vector <- c("Clear","Sunny","Cloudy","Light","Light","Light","Light","Light","Light","Light","Light","Medium","Cloudy","Light","Light","Heavy","Heavy","Heavy","Light","Medium","Heavy","Heavy","Light","Heavy","Heavy","Heavy","Heavy","Heavy","Heavy","Light","Medium","Medium","Light","Heavy","Light","Light","Light","Light","Light","Heavy","Light","Medium","Heavy","Heavy","Heavy")
names(map_vector)<- type_of_weather




ingolstadt_weather = ingolstadt_weather %>% 
  mutate(season = ifelse(month(date) %in% c(12,1,2),"winter",NA)) %>%
  mutate(season = ifelse(month(date) %in% c(3,4,5),"spring",season)) %>%
  mutate(season = ifelse(month(date) %in% c(6,7,8),"summer",season)) %>%
  mutate(season = ifelse(month(date) %in% c(9,10,11),"autumn",season))# %>% dplyr::select(-tsun)




day_description_impact = weatherstack_kelheim_daily %>% pivot_wider(names_from = description,values_from = n)

#remove NAs
day_description_impact[is.na(day_description_impact)] = 0

day_description_impact = day_description_impact %>% pivot_longer(cols = all_of(type_of_weather),names_to = "description",values_to = "value")

day_description_impact = day_description_impact
day_description_impact$description = map_vector[(day_description_impact$description)]

day_description_impact= day_description_impact %>% group_by(date)%>%
  top_n(n = 1,value) %>% group_by(date) %>% top_n(n = 1,description) %>% rename(weather_impact = value)

#####Join the data#####

result_data = demand %>% inner_join(day_description_impact, by = "date") %>% inner_join(ingolstadt_weather,by = "date") %>% inner_join(df_stringency,by = "date") %>% mutate(date = as.Date(date,format = "%Y-%m-%d"))
#Also need to be added weekday
result_data = result_data %>% mutate(wday = as.character(wday(date,week_start = 1)))

#Append holidays
result_data = result_data %>% left_join(df_holidays, by = "date") %>% replace_na(list(isHoliday = FALSE,snow = 0)) #%>% filter(noRides != 0) #%>% filter(date <"2021-07-01")

head(result_data)
```

From following plots we can see strong correlation between day of the week and number of Rides, because we want to simulate common day using MatSim it was decided to filter weekends out of resulting dataset for the model. Also because day doesn't represent weather impact, like holidays that have strong impact on mobility so they are also excluded. First 4 days are excluded, because they have 0 rides due to KeXi service start.

```{r plotting wday and holidays}
wday_plot = ggplotly(ggplot(result_data)+
  geom_boxplot(aes(x = wday,y = noRides)))

holiday_plot = ggplotly(ggplot(result_data)+
  geom_boxplot(aes(x = isHoliday,y = noRides )))

annotations = list( 
  list( 
     x = 0.2,  
    y = 1.0,  
    text = "Weekday",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
     x = 0.75,  
    y = 1.0,  
    text = "Is Holiday",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ))

subplot(wday_plot,holiday_plot) %>% layout(annotations = annotations)
```

```{r}
result_data = result_data %>% filter(wday!=6 & wday!=7,isHoliday == FALSE, noRides!=0)
```

After first data processing it would be helpful to find some dependencies in the data using scatter plots mapped to number of rides.
Here is summary of end dataset 
```{r}
result_data$description = factor(result_data$description)
result_data$season = factor(result_data$season)
summary(result_data)
```

# Finding patterns using graphic approach



```{r noRides~.}

tavg_plot = ggplotly(
  ggplot(result_data)+
    geom_point(aes(y= noRides,x = tavg,colour = season))
  )


pres_plot= ggplotly(
  ggplot(result_data)+
    geom_point(aes(y= noRides,x = pres,colour = season))
  )
prcp_plot = ggplotly(
  ggplot(result_data %>% filter(prcp!=0))+
    geom_point(aes(y= noRides,x = prcp,colour = season))
  )
snow_plot = ggplotly(
  ggplot(result_data %>% filter(snow!=0))+
    geom_point(aes(y= noRides,x = snow,colour = season))
  )

subplot(tavg_plot,pres_plot,prcp_plot,snow_plot,nrows = 2)


```

After looking at these plots we can barely see existing strong dependencies or clusters, but there is one conspicious fact, that only during summer number of rides falls under ~50 while other seasons have more than 50 rides.

# Finding correlations

Correlation coefficients between two variables we calculate using cor() function, later anova oneway test will be used for testing to check categorical variable dependency. 

```{r correlations, warning = FALSE}
best_pred <- result_data %>% ungroup() %>%
  dplyr::select(-noRides,-description ,-date,-season,-noRequests,-avgEuclidianDistance_m,-avgTravelTime_s,-wday) %>%
  map_dbl(cor,y = result_data$noRides) %>%
  #map_dbl(abs) %>%
  sort(decreasing = TRUE) 
print(best_pred)

```

As a result highest absolute score in correlation have temperature(max), snow, pressure(air pressure), stringency(strictness of covid19 policy), weather impact(artificial made variable of highest weather description hours).


Testing of null hypothesis is made for dependent noRides and independent season, description and wday
```{r categorical variables}

season_test = oneway.test(result_data$noRides~result_data$season)$p.value
description_test = oneway.test(result_data$noRides~result_data$description)$p.value
wday_test = oneway.test(result_data$noRides~result_data$wday)$p.value
print(c("season" =season_test,"description" =description_test,"wday" =wday_test))

```

From the p-value we can see that null hypothesis can only be rejected for wday (p.value higher than 0.05), for the description and season we can't make such conclusion. But one thing that should be also tested is dependency between seasons and temperature - `r oneway.test(result_data$tavg~result_data$season)$p.value`. As we can see temperature dependent from season, so in our future model one of the variable should be dropped.

# Building a model

After determination of significant variables we can build different models using different set of variables, one thing remains same - first approach is to build linear model to make predictions, also to check which predictors impact number of rides most. In our approach we will firstly larger model using all relevant variables from previous chapter, then we reduce number of variables taking smaller subset and synchronous to check how model quality changes. 

```{r omega model}
data = result_data

omega_model = lm(noRides ~ tavg+pres+stringency+snow+weather_impact*description,data = data)

summary(omega_model)

```

Low adjusted R^2 doesn't tells that our model doesn't fit the data. Let's take a look on predicted values and residuals.

```{r omega residuals}

colors = c("actual" = "blue","predicted" = "red","residuals" = "gray50","zerorides" = "purple")
model = omega_model
test_data = data %>% add_predictions(model = model) %>% add_residuals(model = model) %>% mutate(error = ifelse(abs(resid)>=20,"extreme","normal"))


ggplotly(ggplot(test_data %>% filter(year(date)>=2020)) +
  geom_line(aes(x = date,y = noRides,color = "actual"))+
  geom_line(aes(x = date,y = pred,color = "predicted"))+
  geom_line(aes(x = date,y = resid,color = "residuals"))+
  geom_ref_line(h = 0)+
  scale_color_manual(values = colors))

```

As we can see from our residuals plot our data has continuous growing trend, that our model doesn't consider, so taking a date variable to model can drastically improve model "quality".

```{r omega date residuals}

omega_date_model = lm(noRides ~ tavg+pres+stringency+snow+weather_impact*description+date+season,data = data)

summary(omega_date_model)

colors = c("actual" = "blue","predicted" = "red","residuals" = "gray50","zerorides" = "purple")
model = omega_date_model
test_data = data %>% add_predictions(model = model) %>% add_residuals(model = model) %>% mutate(error = ifelse(abs(resid)>=20,"extreme","normal"))


ggplotly(ggplot(test_data %>% filter(year(date)>=2020)) +
  geom_line(aes(x = date,y = noRides,color = "actual"))+
  geom_line(aes(x = date,y = pred,color = "predicted"))+
  geom_line(aes(x = date,y = resid,color = "residuals"))+
  geom_ref_line(h = 0)+
  scale_color_manual(values = colors))

```

As expected R^2 squared increased as well residual standard error is decreased to a number of 15,99. ~0.7 of R-squared is relatively high in weather statistics. To check correctness of used approach residuals have to be normally distributed.

Residuals histogram
```{r residual destributions 1}

barplot <- ggplot(test_data, aes(x = resid ))+
  geom_histogram(aes(y = stat(density)),colour="black", fill="white", binwidth=7)+
  geom_density( fill="#FF6666",adjust = 1,alpha = 0.5) +
  ggtitle("Omega model residuals")

ggplotly(barplot)

```

```{r, out.width="100%",warning=FALSE}
normal_dist = fitdistrplus::fitdist(test_data$resid,"norm")
plot(normal_dist)
```
From described plots we can see that our residuals looks normally distributed with skewness to left tail, with some outliers. But simple Shapiro-Wilk test gives p-value of `r shapiro.test(test_data$resid)$p.value` out, so our residuals aren't 

# Reducing a model

Let's take some variables out and see how model performs on the data and F Statistics. Low correlating variables from "Finding correlations" chapter are snow, stringency and weather_impact multiplied by description. Also there is strong correlation between average temperature and a season so our new model should contain only 1 of the predictors, because temperature have most impact on mobility and more large-scaled (season contains only 4 factors) we will take season out from new reduced_model.

```{r reduced model 1 residuals}

reduced_1_model = lm(noRides ~ tavg+pres+stringency+snow+weather_impact*description+date,data = data)

summary(reduced_1_model)

colors = c("actual" = "blue","predicted" = "red","residuals" = "gray50","zerorides" = "purple")
model = reduced_1_model
test_data = data %>% add_predictions(model = model) %>% add_residuals(model = model) %>% mutate(error = ifelse(abs(resid)>=20,"extreme","normal"))


ggplotly(ggplot(test_data %>% filter(year(date)>=2020)) +
  geom_line(aes(x = date,y = noRides,color = "actual"))+
  geom_line(aes(x = date,y = pred,color = "predicted"))+
  geom_line(aes(x = date,y = resid,color = "residuals"))+
  geom_ref_line(h = 0)+
  scale_color_manual(values = colors))

```

As we can see excluding variables from model doesn't make it worse, we can also use anova test function from stats package to check model varable significance. But firstly we would like to get best best model with possible minimum variables. For this purpose we can inspect p-value from model summary, it says how significant separate variable in model predictions, less p-value - more significant. So we would like to exclude snow, and combined term ow weather_impact and description out of model.

```{r reduced model 2 residuals}

reduced_2_model = lm(noRides ~ tavg+pres+stringency+weather_impact+description+date,data = data)

summary(reduced_2_model)

colors = c("actual" = "blue","predicted" = "red","residuals" = "gray50","zerorides" = "purple")
model = reduced_2_model
test_data = data %>% add_predictions(model = model) %>% add_residuals(model = model) %>% mutate(error = ifelse(abs(resid)>=20,"extreme","normal"))


ggplotly(ggplot(test_data %>% filter(year(date)>=2020)) +
  geom_line(aes(x = date,y = noRides,color = "actual"))+
  geom_line(aes(x = date,y = pred,color = "predicted"))+
  geom_line(aes(x = date,y = resid,color = "residuals"))+
  geom_ref_line(h = 0)+
  scale_color_manual(values = colors))

```

As we can see removing snow and separation of weather_impact and description, increased model quality with reducing residual standard error. Anova test between these two models also doesn't show significant difference `r anova(reduced_2_model,reduced_1_model)[["Pr(>F)"]][2]` because p-value is 0.85>>0.05.

After few iterations we get to a model that contains only strictness of the policy restrictions, temperature and data. Represented model explains also explains most of the observations with relatively low residual standard error, but it is minimal for excluding, because anova significance difference between models without any of remaining predictors has p-value<0.05. So we get to our final model, that explains number of KeXi Rides using temperature, covid strictness level and date(positive growing trend). 

```{r reduced model 3 residuals}

reduced_3_model = lm(noRides ~ season+tavg+stringency+date,data = data)

summary(reduced_3_model)

colors = c("actual" = "blue","predicted" = "red","residuals" = "gray50","zerorides" = "purple")
model = reduced_3_model
test_data = data %>% add_predictions(model = model) %>% add_residuals(model = model) %>% mutate(error = ifelse(abs(resid)>=20,"extreme","normal"))


ggplotly(ggplot(test_data %>% filter(year(date)>=2020)) +
  geom_line(aes(x = date,y = noRides,color = "actual"))+
  geom_line(aes(x = date,y = pred,color = "predicted"))+
  geom_line(aes(x = date,y = resid,color = "residuals"))+
  geom_ref_line(h = 0)+
  scale_color_manual(values = colors))

```

After performing a minimal model we also want to check once again how residuals are distributed and some additional statistics metrics.

```{r residual destributions 3}

barplot <- ggplot(test_data, aes(x = resid ))+
  geom_histogram(aes(y = stat(density)),colour="black", fill="white", binwidth=9)+
  geom_density( fill="#FF6666",adjust = 10,alpha = 0.5) +
  ggtitle("Final model residuals")

ggplotly(barplot)

```

We see normal distributed variable with skewness to the legt with 3 outliers at the left (analysis of individual days shows, that extremely low number of rides were when operator collecting data has changed so this is data specific problem that could be ignored or excluded from model)
```{r}
test_data = test_data %>% filter(resid>=-50)
```

```{r, out.width="100%",warning=FALSE}
normal_dist = fitdistrplus::fitdist(test_data$resid,"norm")
plot(normal_dist)
```

```{r residuals verteilung}
fitdistrplus::descdist(test_data$resid)
```

So our residuals are normally distributed :)

Excluding more variables from a model doesn't impove model as well increases residual error enormous, so we can make conclusion, that represented model is minimal fitting possible with variables that have most impact to number of rides by the weather. Remains only to check residuals of fitted model.  

# Conclusion

Performed LINEAR regression analysis shows, that there no strong correlations between number of rides and weather characteristics except of temperature, and temperature is also quite dependent from season, one difference between using season instead of the temperature is that relative temp changes in same season isn't inspected, so it can be that relative fall of the temperature in summer also strongly impacts mobility. Also data was collected during covid-19 pandemie, so stringency-strictness is also significant by building an model. In the end date is used in the model, because there are positive growing trend with using KeXi ride service.

