---
title: "Untitled"
output: html_document
date: "2026-07-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# PRELIM


### libraries
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(lubridate)
library(hms)
library(xml2)
```

### basic paths
```{r}
output_path_base <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-03/output-300it-base-25pct/kelheim-v3.1-25pct-iter_300."

output_path_small <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-03/output-300it-10veh-25pct/kelheim-v3.1-25pct-kexi-iter_300."

output_path_medium <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-03/output-300it-100veh-25pct/kelheim-v3.1-25pct-kexi-iter_300."

output_path_large <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-03/output-300it-1000veh-25pct/kelheim-v3.1-25pct-kexi-iter_300."

```



### read files
```{r}

kelheim_shp <- st_read("/Users/jakob/git/matsim-kelheim/input/shp/lk-kelheim/lk-kelheim.shp")


trips_base <- read_delim(paste0(output_path_base,"output_trips.csv.zst"), delim = ";")
trips_small <- read_delim(paste0(output_path_small,"output_trips.csv.zst"), delim = ";")
trips_medium <- read_delim(paste0(output_path_medium,"output_trips.csv.zst"), delim = ";")
trips_large <- read_delim(paste0(output_path_large,"output_trips.csv.zst"), delim = ";")


persons_base <- read_delim(paste0(output_path_base,"output_persons.csv.zst"),delim = ";")
persons_small <- read_delim(paste0(output_path_small,"output_persons.csv.zst"),delim = ";")
persons_medium <- read_delim(paste0(output_path_medium,"output_persons.csv.zst"),delim = ";")
persons_large <- read_delim(paste0(output_path_large,"output_persons.csv.zst"),delim = ";")

drt_legs_small <- read_delim(paste0(output_path_small,"output_drt_legs_drt.csv"),delim = ";")
drt_legs_medium <- read_delim(paste0(output_path_medium,"output_drt_legs_drt.csv"),delim = ";")
drt_legs_large <- read_delim(paste0(output_path_large,"output_drt_legs_drt.csv"),delim = ";")

occ_small <- read_delim(paste0(output_path_small,"output_occupancy_time_profiles_drt.txt"),delim = ";")
occ_medium <- read_delim(paste0(output_path_medium,"output_occupancy_time_profiles_drt.txt"),delim = ";")
occ_large <- read_delim(paste0(output_path_large,"output_occupancy_time_profiles_drt.txt"),delim = ";")


persons_in_kelheim <- persons_small %>% filter(!is.na(home_x)) %>% st_as_sf(coords = c("home_x","home_y"), crs = 25832) %>% st_filter(kelheim_shp) %>% pull(person)
```
# ANALYSES
## REAL DRT TRIPS
### mode stats 
```{r}
mode_stats_small_trip <- trips_small %>% filter(person %in% persons_in_kelheim) %>% select(main_mode) %>% group_by(main_mode) %>% summarise(n_small = n()) %>% ungroup() %>% mutate(share_10veh = n_small/sum(n_small))

mode_stats_medium_trip <- trips_medium %>% filter(person %in% persons_in_kelheim) %>% select(main_mode) %>% group_by(main_mode) %>% summarise(n_medium = n()) %>% ungroup() %>% mutate(share_100veh = n_medium/sum(n_medium)) 
mode_stats_large_trip <- trips_large %>% filter(person %in% persons_in_kelheim) %>% select(main_mode) %>% group_by(main_mode) %>% summarise(n_large = n()) %>% ungroup() %>% mutate(share_1000veh = n_large/sum(n_large))

mode_stats_small_trip %>% left_join(mode_stats_medium_trip) %>% left_join(mode_stats_large_trip) 

```

### score stats
```{r}
# Whole Population
person_score_joined <- persons_base %>% select(person,executed_score_base = executed_score) %>% left_join(persons_small %>% select(person,executed_score_small = executed_score)) %>%  left_join(persons_medium %>% select(person,executed_score_medium = executed_score)) %>% left_join(persons_large %>% select(person,executed_score_large = executed_score))
person_score_joined

person_score_joined %>%
  summarise(
    across(
      starts_with("exe"),
      list(mean = mean, sum = sum)
    )
  )

# Affected People
drt_users_small <- trips_small %>% filter(main_mode == "drt") %>% pull(person) %>% unique()
drt_users_medium <- trips_medium %>% filter(main_mode == "drt") %>% pull(person) %>% unique()
drt_users_large <- trips_large %>% filter(main_mode == "drt") %>% pull(person) %>% unique()



person_score_joined %>% filter(person %in% drt_users_small) %>% mutate(diff = executed_score_small - executed_score_base) %>% summarise(diff = mean(diff))

person_score_joined %>% filter(person %in% drt_users_medium) %>% mutate(diff = executed_score_medium - executed_score_base) %>% summarise(diff = mean(diff))

person_score_joined %>% filter(person %in% drt_users_large) %>% mutate(diff = executed_score_large - executed_score_base) %>% summarise(diff = mean(diff))

```

### requests


```{r}
bind_rows(drt_legs_small %>% mutate(vehicles = 10), drt_legs_medium %>% mutate(vehicles = 100), drt_legs_large %>% mutate(vehicles = 1000)) %>%
  filter(!startsWith(personId,"COMP")) %>%
  mutate(time = hms::as_hms(submissionTime)) %>%
  mutate(hour = hour(time)) %>% 
  mutate(vehicles = as.factor(vehicles)) %>% 
  group_by(hour, vehicles) %>% count() %>% ggplot() + geom_line(aes(hour,n, col = vehicles)) + theme_minimal(base_size = 14) + labs(x = "Hour", y = "Requests")
```


### wait time and detour factor

```{r}

drt_legs_combined <- bind_rows(drt_legs_small %>% mutate(vehicles = 10), drt_legs_medium %>% mutate(vehicles = 100), drt_legs_large %>% mutate(vehicles = 1000)) %>%
  filter(!startsWith(personId,"COMP")) %>% 
  group_by(vehicles) %>%
  summarise(waitTime = mean(waitTime),inVehicleRideTime = mean(inVehicleRideTime),directRideTime = mean(directRideTime)) %>%
  mutate(detourFactor = inVehicleRideTime/directRideTime) %>% 
  mutate(vehicles = as.factor(vehicles))


ggplot(drt_legs_combined) + geom_col(aes(vehicles, waitTime)) + theme_minimal(base_size = 14)
ggplot(drt_legs_combined) + geom_col(aes(vehicles, directRideTime))+ theme_minimal(base_size = 14)
ggplot(drt_legs_combined) + geom_col(aes(vehicles, inVehicleRideTime))+ theme_minimal(base_size = 14)
ggplot(drt_legs_combined) + geom_col(aes(vehicles, detourFactor))+ theme_minimal(base_size = 14)


# drt_legs_small %>% filter(!startsWith(personId,"COMP")) %>% summarise(waitTime = mean(waitTime),inVehicleRideTime = mean(inVehicleRideTime),directRideTime = mean(directRideTime)) %>% mutate(detourFactor = inVehicleRideTime/directRideTime)
# 
# drt_legs_medium %>% filter(!startsWith(personId,"COMP")) %>% summarise(waitTime = mean(waitTime),inVehicleRideTime = mean(inVehicleRideTime),directRideTime = mean(directRideTime)) %>% mutate(detourFactor = inVehicleRideTime/directRideTime)
# 
# drt_legs_large %>% filter(!startsWith(personId,"COMP")) %>% summarise(waitTime = mean(waitTime),inVehicleRideTime = mean(inVehicleRideTime),directRideTime = mean(directRideTime)) %>% mutate(detourFactor = inVehicleRideTime/directRideTime)
```

### occupancy
```{r}

occ_small %>% mutate(fleet_size = 10) %>% 
  bind_rows(occ_medium %>% mutate(fleet_size = 100)) %>% 
  bind_rows(occ_large %>% mutate(fleet_size = 1000)) %>% 
  filter(time <= hm("24:00")) %>% 
  pivot_longer(-c(time,fleet_size), names_to = "status", values_to = "n") %>% 
  # mutate(status = factor(status, levels = c("STAY","0 pax", "1 pax", "2 pax", "3 pax", "4 pax"))) %>% 
  ggplot() + 
  scale_x_time(minor_breaks = "1 hour", date_breaks = "3 hours",date_labels = "%H") +
  geom_area(aes(time, n, fill = status)) + 
  scale_fill_manual(values = c(
    "STAY" = "gray",
    "0 pax" = "white",
    "1 pax" = "yellow",
    "2 pax" = "orange",
    "3 pax" = "red",
    "4 pax" = "purple"
  )) +
  facet_grid(~fleet_size)+
  theme_minimal(base_size = 16) + 
  labs(x = "Time", y = "Number of Vehicles")
```

## POTENTIAL DRT TRIPS
### probes

The way we did it before was biased. We assumed potential trip has the same LOS as the average "realized" trips. However, in MATSim, only the people who had good DRT LOS would actually take it. The people with long wait times or detour factors wouldn't actually take the DRT, thus keeping the average LOS of realized trips relatively high. Our new methodology is way better; it essentially checks the LOS of also unrealized trips.

read drt stops
```{r}
doc <- read_xml("/Users/jakob/git/matsim-kelheim/input/drt_stops_landkreis.xml")

stops <- xml_find_all(doc, ".//stopFacility")

drt_stops <- tibble(
  id = as.numeric(xml_attr(stops, "id")),
  x = as.numeric(xml_attr(stops, "x")),
  y = as.numeric(xml_attr(stops, "y"))
) %>%
  st_as_sf(coords = c("x","y"), crs = 25832)
```


Prepare data

```{r}
# data <- read_delim("/Users/jakob/git/matsim-kelheim/output/output-kelheim-v3.0-1pct-iter_1-plans_kelheim-v3-1-1pct-kexi-iter_300-output_plans/kelheim-v3.1-1pct-kexi-iter_1-plans_kelheim-v3-1-1pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 


# train station 
# probe_small <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-10-c-train-only-0it-5min/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 
# probe_medium <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-100-c-train-only-0it-5min/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 
# probe_large <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-1000-c-train-only-0it-5min/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 

# supermarket 
probe_small <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-10-d-supermarket-only-0it-hourly/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 
probe_medium <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-100-d-supermarket-only-0it-hourly/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 
probe_large <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-1000-d-supermarket-only-0it-hourly/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 


probe_combined <- bind_rows(probe_small %>% mutate(vehicles = 10), probe_medium %>% mutate(vehicles = 100),probe_large %>% mutate(vehicles = 1000)) %>% 
  mutate(time = hms::as_hms(time)) %>% 
  mutate(hour = hour(time)) %>% 
  group_by(hour,vehicles,originStop,destinationStop) %>% 
  summarise(waitTime = mean(waitTime), directRideTime = mean(directRideTime), rideTimeWithDetour = mean(rideTimeWithDetour),directRideDistance=mean(directRideDistance), rideDistanceWithDetour = mean(rideDistanceWithDetour), detourFactor = mean(detourFactor)) %>% 
  mutate(hour = as.factor(hour))

probe_combined 
```


```{r}
probe_combined %>%
  ggplot() + 
  geom_boxplot(aes(hour,waitTime)) +
  facet_grid(~vehicles) +
  theme_minimal(base_size = 14) + 
  labs(x = "Time", y = "Wait Time (sec)")

probe_combined %>% 
  ggplot() + 
  geom_boxplot(aes(hour,detourFactor)) +
  facet_grid(~vehicles) +
  theme_minimal(base_size = 14) + 
  labs(x = "Time", y = "Detour Factor")


tmap_mode("plot")
probe_combined %>% 
  filter(hour == 18) %>% 
  group_by(originStop, vehicles) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor)) %>% 
  left_join(drt_stops, by = c("originStop" = "id")) %>% 
  st_as_sf() %>% 
  # group_by(originZone, vehicles) %>% 
  # summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor), originX = first(originX), originY = first(originY)) %>%
  # st_as_sf(coords = c("originX","originY"), crs=25832) %>%
  tm_shape() + 
  tm_basemap("Esri.WorldGrayCanvas") +
  tm_dots(fill = "detourFactor", size = .4) + 
  tm_title("Detour Factor by Origin") + 
  tm_facets(by = "vehicles")


probe_combined %>% 
 filter(hour == 18) %>% 
  group_by(destinationStop, vehicles) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor)) %>% 
  left_join(drt_stops, by = c("destinationStop" = "id")) %>% 
  st_as_sf() %>% 
  tm_shape() + 
  tm_basemap("Esri.WorldGrayCanvas") +
  tm_dots(fill = "detourFactor", size = .4) + 
  tm_title("Detour Factor by Destination") + 
  tm_facets(by = "vehicles")

probe_combined %>% 
  filter(hour == 18) %>%
  group_by(originStop, vehicles) %>% 
  left_join(drt_stops, by = c("originStop" = "id")) %>% 
  st_as_sf() %>% 
  summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor)) %>% 
  tm_shape() + 
  tm_basemap("Esri.WorldGrayCanvas") +
  tm_dots(fill = "waitTime", size = .4, fill.scale = tm_scale_intervals(style = "quantile",n = 10)) +
  tm_title("Wait Time by Origin")+ 
  tm_facets(by = "vehicles")
```
### accessibility
```{r}

#train station
# acc_small <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-10-b-10veh/analysis/accessibility/train_station/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>% 
#   mutate(time = as_hms(time)) %>% mutate(vehicles = 10)
# acc_medium <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-10-b-100veh/analysis/accessibility/train_station/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>% 
#   mutate(time = as_hms(time))%>% mutate(vehicles = 100)
# acc_large <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-10-b-1000veh/analysis/accessibility/train_station/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>% 
#   mutate(time = as_hms(time))%>% mutate(vehicles = 1000)

#supermarket
acc_small <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-12-supermarket-10veh/analysis/accessibility/supermarket/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>% 
  mutate(time = as_hms(time)) %>% mutate(vehicles = 10)
acc_medium <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-12-supermarket-100veh/analysis/accessibility/supermarket/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>% 
  mutate(time = as_hms(time))%>% mutate(vehicles = 100)
acc_large <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-12-supermarket-1000veh/analysis/accessibility/supermarket/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>% 
  mutate(time = as_hms(time))%>% mutate(vehicles = 1000)

acc_combined <- bind_rows(acc_small,acc_medium,acc_large)



acc_combined %>% 
  mutate(hour = hour(time), minute = minute(time), vehicles = as.factor(vehicles)) %>%
  mutate(population = population * 4) %>% 
  mutate(population_density = population / (0.5*0.5)) %>% 
  # filter(minute == 0) %>% 
# filter(time == hm("8:00")) %>% arrange(id) %>% 
  st_drop_geometry() %>% 
  group_by(vehicles,hour) %>% 
  summarise(estimatedDrt_accessibility = sum(estimatedDrt_accessibility*population)/sum(population)) %>% #teleportedWalk_accessibility = mean(teleportedWalk_accessibility), 
  ggplot() +
  geom_col(aes(hour,estimatedDrt_accessibility)) + # fill = hour, position = "dodge"
  facet_grid(~vehicles) +
  labs(x = "Hour", y = "DRT Accessibility (utils)", title = "DRT Accessibility Averaged over Hour for Three Fleet Sizes") +
  theme_minimal(base_size = 14)


```

```{r}

```


#XXX
```{r}
persons_base


# 25,184 adult agents living in Kelheim
persons_base %>% filter(person %in% persons_in_kelheim) %>% filter(age >=18)

# 5,622  immobile adult agents living in kelheim = 22.3%. In the MiD it is 14%
persons_base %>% filter(person %in% persons_in_kelheim) %>% filter(age >=18) %>% left_join(trips_base %>% group_by(person) %>% count()) %>% filter(is.na(n))


# all kelheim adult agents have car availability
persons_base %>% filter(age >=18) %>% group_by(carAvail) %>% count() %>% ungroup() %>% mutate(share = n/sum(n)*100.)




```


# DEPRECATED 
mode stats file
```{r}
mode_stats_small <- read_delim(paste0(output_path_small,"modestats.csv"),delim = ";") %>% 
  pivot_longer(-iteration, names_to = "mode",values_to = "share") %>%  
  filter(mode!="freight") 

mode_stats_small%>% 
  ggplot() + geom_line(aes(iteration,share, col = mode))

mode_stats_large <- read_delim(paste0(output_path_large,"modestats.csv"),delim = ";") %>% 
  pivot_longer(-iteration, names_to = "mode",values_to = "share") %>%  
  filter(mode!="freight") 

mode_stats_large%>% 
  ggplot() + geom_line(aes(iteration,share, col = mode))



mode_stats_small %>% filter(iteration == 300) %>% left_join(mode_stats_large %>% filter(iteration == 300), by = c("iteration","mode")) %>% mutate(diff = share.y - share.x)


# who is using drt - distributed around study area 
# trips_small %>% filter(main_mode == "drt") %>% select(person) %>% left_join(persons_raw) %>% st_as_sf(coords = c("home_x","home_y"), crs = 25832) %>% tm_shape()+ tm_dots()


```


wait time & detour factor 
```{r}
data <- read_delim(paste0(output_path_small,"drt_service_quality_probes.csv.gz"), delim = ";")

data %>% mutate(time = as.factor(time / 3600)) %>% ggplot() + geom_boxplot(aes(time,waitTime))

tmap_mode("view")
data %>% filter(time == 16*3600) %>% group_by(originZone) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor), originX = first(originX), originY = first(originY)) %>% st_as_sf(coords = c("originX","originY"), crs=25832) %>% tm_shape() + tm_dots(fill = "detourFactor", size = 2)

data %>% filter(time == 16*3600) %>% group_by(originZone) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor), originX = first(originX), originY = first(originY)) %>% st_as_sf(coords = c("originX","originY"), crs=25832) %>% tm_shape() + tm_dots(fill = "waitTime", size = 2)


```


```{r}
read_delim("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-03/output-300it-10veh-1pct-500m/kelheim-v3.1-1pct-kexi-iter_300.drt_service_quality_probes.csv.gz", delim = ";") %>% filter(time == 21600) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor))

read_delim("/Users/jakob/git/matsim-kelheim/output/output-kelheim-v3.0-1pct-iter_1-plans_kelheim-v3-1-1pct-kexi-iter_300-output_plans/kelheim-v3.1-1pct-kexi-iter_1-plans_kelheim-v3-1-1pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") %>% view() 

summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor))
```

