---
title: "Untitled"
output: html_document
date: "2026-07-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# PRELIM


### libraries
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(lubridate)
library(hms)
library(xml2)
```

### basic paths
```{r}
library(tibble)

df_full <- tribble(
  ~fleet_size, ~demand,      ~path,
  NA_integer_, "base","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/base__demand-base__seed-4711/kelheim-v3.1-25pct-iter_300.",
  10L,"base","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-10__seed-4711/kelheim-v3.1-25pct-kexi-iter_300.",
  50L,"base","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-50__seed-4711/kelheim-v3.1-25pct-kexi-iter_300.",
  100L,"base","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-100__seed-4711/kelheim-v3.1-25pct-kexi-iter_300.",
  500L,"base","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-500__seed-4711/kelheim-v3.1-25pct-kexi-iter_300.",
  1000L,"base","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-1000__seed-4711/kelheim-v3.1-25pct-kexi-iter_300.",
  NA_integer_, "augmented","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/base__demand-counterfactual__seed-4711/kelheim-v3.1-25pct-iter_300-plans_counterfactual-25pct-plans.",
  10L,"augmented","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-10__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.",
  50L,"augmented","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-50__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.",
  100L,"augmented","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-100__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.",
  500L,"augmented","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-500__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.",
  1000L,"augmented","/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-1000__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans."
)
```


```{r}

library(purrr)

library(sf)

df_full2 <- df_full %>%
  mutate(
    run_id = paste(demand, if_else(is.na(fleet_size), "baseline", paste0("fleet-", fleet_size)), sep = "__"),
    fleet_label = if_else(is.na(fleet_size), "baseline", paste0("fleet-", fleet_size))
  ) %>%
  mutate(
    trips = purrr::map(
      path,
      ~ readr::read_delim(
        paste0(.x, "output_trips.csv.zst"),
        delim = ";",
        show_col_types = FALSE
      )
    ),

    persons = purrr::map(
      path,
      ~ readr::read_delim(
        paste0(.x, "output_persons.csv.zst"),
        delim = ";",
        show_col_types = FALSE
      )
    ),

    drt_legs = purrr::map2(
      path,
      fleet_size,
      ~ if (!is.na(.y)) {
        readr::read_delim(
          paste0(.x, "output_drt_legs_drt.csv"),
          delim = ";",
          show_col_types = FALSE
        )
      } else {
        NULL
      }
    ),

    occupancy = purrr::map2(
      path,
      fleet_size,
      ~ if (!is.na(.y)) {
        readr::read_delim(
          paste0(.x, "output_occupancy_time_profiles_drt.txt"),
          delim = ";",
          show_col_types = FALSE
        )
      } else {
        NULL
      }
    )
  )
```


```{r}
# 
# # BASE DEMAND
# output_path_base <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/base__demand-base__seed-4711/kelheim-v3.1-25pct-iter_300."
# output_path_10 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-10__seed-4711/kelheim-v3.1-25pct-kexi-iter_300."
# output_path_50 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-50__seed-4711/kelheim-v3.1-25pct-kexi-iter_300."
# output_path_100 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-100__seed-4711/kelheim-v3.1-25pct-kexi-iter_300."
# output_path_500 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-500__seed-4711/kelheim-v3.1-25pct-kexi-iter_300."
# output_path_1000 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-base__prebooking-off__fleet-1000__seed-4711/kelheim-v3.1-25pct-kexi-iter_300."
# 
# output_path_small = output_path_10
# output_path_medium = output_path_100
# output_path_large = output_path_1000
# 
# # AUGMENTED DEMAND
# 
# # output_path_base <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/base__demand-counterfactual__seed-4711/kelheim-v3.1-25pct-iter_300-plans_counterfactual-25pct-plans."
# # output_path_10 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-10__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.annealingRates.txt"
# # output_path_50 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-50__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.annealingRates.txt"
# # output_path_100 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-100__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.annealingRates.txt"
# # output_path_500 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-500__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.annealingRates.txt"
# # output_path_1000 <- "/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-15-counterfactual-fleet/policy__demand-counterfactual__prebooking-off__fleet-1000__seed-4711/kelheim-v3.1-25pct-kexi-iter_300-plans_counterfactual-25pct-plans.annealingRates.txt"

```



### read files
```{r}

kelheim_shp <- st_read("/Users/jakob/git/matsim-kelheim/input/shp/lk-kelheim/lk-kelheim.shp")

df_full2 <- df_full2 %>%
  mutate(
    persons_in_kelheim = purrr::map(
      persons,
      ~ .x %>%
        filter(!is.na(home_x), !is.na(home_y)) %>%
        st_as_sf(coords = c("home_x", "home_y"), crs = 25832) %>%
        st_filter(kelheim_shp) %>%
        st_drop_geometry() %>%
        pull(person)
    )
  )
```
# ANALYSES
## REAL DRT TRIPS
### mode stats 
```{r}
mode_stats <- purrr::pmap_dfr(
  list(df_full2$trips, df_full2$persons_in_kelheim, df_full2$run_id, df_full2$demand, df_full2$fleet_label),
  function(trips, kelheim_persons, run_id, demand, fleet_label) {
    trips %>%
      filter(person %in% kelheim_persons) %>%
      count(main_mode, name = "n") %>%
      mutate(
        share = n / sum(n) * 100,
        run_id = run_id,
        demand = demand,
        fleet = fleet_label
      )
  }
)

mode_stats %>% filter(main_mode == "ride")

```

### score stats
```{r}
person_scores <- purrr::map2_dfr(
  df_full2$persons,
  df_full2$run_id,
  ~ .x %>% select(person, executed_score) %>% mutate(run_id = .y)
)

score_stats <- person_scores %>%
  left_join(df_full2 %>% select(run_id, demand, fleet = fleet_label), by = "run_id") %>%
  group_by(run_id, demand, fleet) %>%
  summarise(
    mean_score = mean(executed_score, na.rm = TRUE),
    total_score = sum(executed_score, na.rm = TRUE),
    .groups = "drop"
  )

baseline_scores <- person_scores %>%
  left_join(df_full2 %>% select(run_id, demand, fleet = fleet_label), by = "run_id") %>%
  filter(fleet == "baseline") %>%
  select(demand, person, baseline_score = executed_score)

drt_score_differences <- purrr::pmap_dfr(
  list(df_full2$trips, df_full2$run_id, df_full2$demand, df_full2$fleet_label),
  function(trips, current_run_id, current_demand, current_fleet) {
    if (current_fleet == "baseline") return(tibble())

    trips %>%
      filter(main_mode == "drt") %>%
      distinct(person) %>%
      left_join(person_scores %>% filter(.data$run_id == current_run_id), by = "person") %>%
      left_join(baseline_scores %>% filter(.data$demand == current_demand), by = "person") %>%
      summarise(
        run_id = current_run_id,
        demand = current_demand,
        fleet = current_fleet,
        n_drt_users = n(),
        mean_score_difference = mean(executed_score - baseline_score, na.rm = TRUE)
      )
  }
)

score_stats
drt_score_differences

```

### requests


```{r}
drt_legs_all <- purrr::map2_dfr(
  df_full2$drt_legs,
  df_full2$run_id,
  ~ if (is.null(.x)) tibble() else .x %>% mutate(run_id = .y)
) %>%
  left_join(df_full2 %>% select(run_id, demand, fleet = fleet_label), by = "run_id") %>%
  filter(!startsWith(personId, "COMP")) %>%
  mutate(time = hms::as_hms(submissionTime), hour = hour(time))

request_counts <- drt_legs_all %>%
  count(demand, fleet, hour, name = "requests")

ggplot(request_counts) +
  geom_line(aes(hour, requests, colour = fleet, group = fleet)) +
  facet_grid(~ demand) +
  theme_minimal(base_size = 14) +
  labs(x = "Hour", y = "Requests")
```


### wait time and detour factor

```{r}

drt_legs_combined <- drt_legs_all %>%
  group_by(demand, fleet) %>%
  summarise(waitTime = mean(waitTime, na.rm = TRUE), inVehicleRideTime = mean(inVehicleRideTime, na.rm = TRUE), directRideTime = mean(directRideTime, na.rm = TRUE), .groups = "drop") %>%
  mutate(detourFactor = inVehicleRideTime/directRideTime)

drt_legs_combined


for (metric in c("waitTime", "directRideTime", "inVehicleRideTime", "detourFactor")) {
  print(
    ggplot(drt_legs_combined) +
      geom_col(aes(fleet, .data[[metric]])) +
      facet_grid(~ demand) +
      theme_minimal(base_size = 14) +
      labs(x = "Fleet", y = metric)
  )
}
```

### occupancy
```{r}

occupancy_all <- purrr::map2_dfr(
  df_full2$occupancy,
  df_full2$run_id,
  ~ if (is.null(.x)) tibble() else .x %>% mutate(run_id = .y)
) %>%
  left_join(df_full2 %>% select(run_id, demand, fleet = fleet_label), by = "run_id")

occupancy_all %>%
  mutate(fleet = factor(fleet, levels = unique(df_full2$fleet_label))) %>%
  filter(time <= hm("24:00")) %>% 
  pivot_longer(-c(time, run_id, demand, fleet), names_to = "status", values_to = "n") %>%
  # mutate(status = factor(status, levels = c("STAY","0 pax", "1 pax", "2 pax", "3 pax", "4 pax"))) %>% 
  ggplot() + 
  scale_x_time(minor_breaks = "1 hour", date_breaks = "3 hours",date_labels = "%H") +
  geom_area(aes(time, n, fill = status)) + 
  scale_fill_manual(values = c(
    "STAY" = "gray",
    "0 pax" = "white",
    "1 pax" = "yellow",
    "2 pax" = "orange",
    "3 pax" = "red",
    "4 pax" = "purple"
  )) +
  facet_grid(demand ~ fleet)+
  theme_minimal(base_size = 16) + 
  labs(x = "Time", y = "Number of Vehicles")
```

## POTENTIAL DRT TRIPS
### probes

The way we did it before was biased. We assumed potential trip has the same LOS as the average "realized" trips. However, in MATSim, only the people who had good DRT LOS would actually take it. The people with long wait times or detour factors wouldn't actually take the DRT, thus keeping the average LOS of realized trips relatively high. Our new methodology is way better; it essentially checks the LOS of also unrealized trips.

read drt stops
```{r}
doc <- read_xml("/Users/jakob/git/matsim-kelheim/input/v3.1/expanded-service-area/drt_stops_landkreis.xml")

stops <- xml_find_all(doc, ".//stopFacility")

drt_stops <- tibble(
  id = as.numeric(xml_attr(stops, "id")),
  x = as.numeric(xml_attr(stops, "x")),
  y = as.numeric(xml_attr(stops, "y"))
) %>%
  st_as_sf(coords = c("x","y"), crs = 25832)
```


Prepare data

```{r}
# data <- read_delim("/Users/jakob/git/matsim-kelheim/output/output-kelheim-v3.0-1pct-iter_1-plans_kelheim-v3-1-1pct-kexi-iter_300-output_plans/kelheim-v3.1-1pct-kexi-iter_1-plans_kelheim-v3-1-1pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 


# train station 
# probe_small <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-10-c-train-only-0it-5min/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 
# probe_medium <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-100-c-train-only-0it-5min/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 
# probe_large <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-1000-c-train-only-0it-5min/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 

# supermarket 
# probe_small <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-10-d-supermarket-only-0it-hourly/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 
# probe_medium <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-100-d-supermarket-only-0it-hourly/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 
# probe_large <- read_delim("/Users/jakob/git/matsim-kelheim/output-probe-1000-d-supermarket-only-0it-hourly/kelheim-v3.1-25pct-kexi-iter_0-plans_kelheim-v3-1-25pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") 

probe_small <- read_delim(paste0(output_path_small,"drt_service_quality_probes.csv.gz"), delim = ";")
probe_medium <- read_delim(paste0(output_path_medium,"drt_service_quality_probes.csv.gz"), delim = ";")
probe_large <- read_delim(paste0(output_path_large,"drt_service_quality_probes.csv.gz"), delim = ";")


probe_combined <- bind_rows(probe_small %>% mutate(fleet = "small"), probe_medium %>% mutate(fleet = "medium"),probe_large %>% mutate(fleet = "large")) %>% 
  mutate(fleet = factor(fleet, levels = c("small","medium","large"))) %>% 
  mutate(time = hms::as_hms(time)) %>% 
  mutate(hour = hour(time)) %>% 
  group_by(hour,fleet,originStop,destinationStop) %>% 
  summarise(waitTime = mean(waitTime), directRideTime = mean(directRideTime), rideTimeWithDetour = mean(rideTimeWithDetour),directRideDistance=mean(directRideDistance), rideDistanceWithDetour = mean(rideDistanceWithDetour), detourFactor = mean(detourFactor)) %>% 
  mutate(hour = as.factor(hour))

probe_combined 
```


```{r}
probe_combined %>%
  ggplot() + 
  geom_boxplot(aes(hour,waitTime)) +
  facet_grid(~fleet) +
  theme_minimal(base_size = 14) + 
  labs(x = "Time", y = "Wait Time (sec)")

probe_combined %>% 
  ggplot() + 
  geom_boxplot(aes(hour, detourFactor)) +
  facet_grid(~fleet) +
  theme_minimal(base_size = 14) + 
  labs(x = "Time", y = "Detour Factor")


tmap_mode("plot")
probe_combined %>% 
  filter(hour == 18) %>% 
  group_by(originStop, fleet) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor)) %>% 
  left_join(drt_stops, by = c("originStop" = "id")) %>% 
  st_as_sf() %>% 
  # group_by(originZone, vehicles) %>% 
  # summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor), originX = first(originX), originY = first(originY)) %>%
  # st_as_sf(coords = c("originX","originY"), crs=25832) %>%
  tm_shape() + 
  tm_basemap("Esri.WorldGrayCanvas") +
  tm_dots(fill = "detourFactor", size = .4) + 
  tm_title("Detour Factor by Origin") + 
  tm_facets(by = "fleet")


probe_combined %>% 
 filter(hour == 18) %>% 
  group_by(destinationStop, fleet) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor)) %>% 
  left_join(drt_stops, by = c("destinationStop" = "id")) %>% 
  st_as_sf() %>% 
  tm_shape() + 
  tm_basemap("Esri.WorldGrayCanvas") +
  tm_dots(fill = "detourFactor", size = .4) + 
  tm_title("Detour Factor by Destination") + 
  tm_facets(by = "fleet")

probe_combined %>% 
  filter(hour == 18) %>%
  group_by(originStop, fleet) %>% 
  left_join(drt_stops, by = c("originStop" = "id")) %>% 
  st_as_sf() %>% 
  summarise(`Wait Time (min)` = mean(waitTime)/60, detourFactor = mean(detourFactor)) %>% 
  tm_shape() + 
  tm_basemap("Esri.WorldGrayCanvas") +
  tm_dots(fill = "Wait Time (min)", size = .4, fill.scale = tm_scale_intervals(style = "fixed",breaks = c(0,2,4,6,8,10,Inf))) +
  tm_title("Wait Time by Origin at 18:00")+ 
  tm_place_legends_right() +
  tm_facets(by = "fleet")


probe_combined %>% 
  filter(fleet == "small") %>%
  filter(hour %in% c("6","8","10","12","14","16","18","20")) %>% 
  group_by(originStop, hour) %>% 
  left_join(drt_stops, by = c("originStop" = "id")) %>% 
  st_as_sf() %>% 
  summarise(`Wait Time (min)` = mean(waitTime)/60, detourFactor = mean(detourFactor)) %>% 
  tm_shape() + 
  tm_basemap("Esri.WorldGrayCanvas") +
  tm_dots(fill = "Wait Time (min)", size = .4, fill.scale = tm_scale_intervals(style = "fixed",breaks = c(seq(0,20,2),Inf))) +
  tm_title("Wait Time by Origin, Small Fleet Size")+ 
  tm_place_legends_right() +
  tm_facets(by = "hour",nrow = 2, ncol = 4)

```
### accessibility

TRAIN STATION
```{r}

acc_old <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/kelheim-v3/analysis/accessibility/train_station/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>%
  mutate(time = as_hms(time)) %>% mutate(fleet = "old")
acc_small <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-14/demand-base__prebooking-off__fleet-100__seed-4711/analysis/accessibility/train_station/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>%
  mutate(time = as_hms(time)) %>% mutate(fleet = "small")
acc_medium <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-14/demand-base__prebooking-off__fleet-250__seed-4711/analysis/accessibility/train_station/accessibilities_simwrapper.csv") %>%
  mutate(time = as_hms(time))%>% mutate(fleet = "medium")
acc_large <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-14/demand-base__prebooking-off__fleet-1000__seed-4711/analysis/accessibility/train_station/accessibilities_simwrapper.csv") %>%
  mutate(time = as_hms(time))%>% mutate(fleet = "large")

acc_combined <- bind_rows(acc_old,acc_small,acc_medium,acc_large)



acc_combined %>% 
  mutate(hour = hour(time), minute = minute(time)) %>%
  mutate(population = population * 4) %>% 
  mutate(population_density = population / (0.5*0.5)) %>% 
  mutate(fleet = factor(fleet, levels = c("old","small","medium","large"))) %>% 
  # filter(minute == 0) %>% 
# filter(time == hm("8:00")) %>% arrange(id) %>% 
  st_drop_geometry() %>% 
  group_by(fleet,hour) %>% 
  summarise(estimatedDrt_accessibility = sum(estimatedDrt_accessibility*population)/sum(population)) %>% #teleportedWalk_accessibility = mean(teleportedWalk_accessibility), 
  ggplot() +
  geom_col(aes(hour,estimatedDrt_accessibility)) + # fill = hour, position = "dodge"
  facet_grid(~fleet) +
  labs(x = "Hour", y = "DRT Accessibility (utils)", title = "DRT Accessibility Averaged over Hour for Three Fleet Sizes (TRAIN)") +
  theme_minimal(base_size = 14)


acc_combined %>% 
  mutate(hour = hour(time), minute = minute(time)) %>%
  mutate(population = population * 4) %>% 
  mutate(population_density = population / (0.5*0.5)) %>% 
  mutate(fleet = factor(fleet, levels = c("old","small","medium","large"))) %>% 
  # filter(minute == 0) %>% 
# filter(time == hm("8:00")) %>% arrange(id) %>% 
  st_drop_geometry() %>% 
  filter(hour >=7 & hour <=10) %>% 
  group_by(fleet) %>% 
  summarise(estimatedDrt_accessibility = sum(estimatedDrt_accessibility*population)/sum(population), car_accessibility = sum(car_accessibility * population)/sum(population), pt_accessibility = sum(pt_accessibility * population)/sum(population), teleportedWalk_accessibility = sum(teleportedWalk_accessibility * population)/sum(population)) %>%
    pivot_longer(ends_with("accessibility"), names_to = "mode", values_to = "accessibility") %>% 
mutate(
  mode = case_when(
    mode == "teleportedWalk_accessibility" ~ "walk",
    mode == "estimatedDrt_accessibility"   ~ "drt",
    mode == "pt_accessibility"             ~ "pt",
    mode == "car_accessibility"            ~ "car",
    TRUE ~ mode
  )
) %>% 
  ggplot() +
  geom_col(aes(fleet,accessibility)) + # fill = hour, position = "dodge"
  facet_grid(~mode) +
  labs(x = "Hour", y = "DRT Accessibility (utils)", title = "Accessibility 7-10AM to Train Station") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))






```
SUPERMARKET
```{r}
acc_old <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/kelheim-v3/analysis/accessibility/supermarket/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>%
  mutate(time = as_hms(time)) %>% mutate(fleet = "old")

acc_small <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-14/demand-base__prebooking-off__fleet-100__seed-4711/analysis/accessibility/supermarket/accessibilities_simwrapper.csv") %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>%
  mutate(time = as_hms(time)) %>% mutate(fleet = "small")
acc_medium <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-14/demand-base__prebooking-off__fleet-250__seed-4711/analysis/accessibility/supermarket/accessibilities_simwrapper.csv") %>%
  mutate(time = as_hms(time))%>% mutate(fleet = "medium")
acc_large <- read_csv("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-14/demand-base__prebooking-off__fleet-1000__seed-4711/analysis/accessibility/supermarket/accessibilities_simwrapper.csv") %>%
  mutate(time = as_hms(time))%>% mutate(fleet = "large")

acc_combined <- bind_rows(acc_old,acc_small,acc_medium,acc_large)



acc_combined %>% 
  mutate(hour = hour(time), minute = minute(time)) %>%
  mutate(population = population * 4) %>% 
  mutate(population_density = population / (0.5*0.5)) %>% 
  mutate(fleet = factor(fleet, levels = c("old","small","medium","large"))) %>% 
  # filter(minute == 0) %>% 
# filter(time == hm("8:00")) %>% arrange(id) %>% 
  st_drop_geometry() %>% 
  group_by(fleet,hour) %>% 
  summarise(estimatedDrt_accessibility = sum(estimatedDrt_accessibility*population)/sum(population)) %>% #teleportedWalk_accessibility = mean(teleportedWalk_accessibility), 
  ggplot() +
  geom_col(aes(hour,estimatedDrt_accessibility)) + # fill = hour, position = "dodge"
  facet_grid(~fleet) +
  labs(x = "Hour", y = "DRT Accessibility (utils)", title = "DRT Accessibility Averaged over Hour for Three Fleet Sizes (SUPERMARKET)") +
  theme_minimal(base_size = 14)


acc_combined %>% 
  mutate(hour = hour(time), minute = minute(time)) %>%
  mutate(population = population * 4) %>% 
  mutate(population_density = population / (0.5*0.5)) %>% 
  mutate(fleet = factor(fleet, levels = c("old","small","medium","large"))) %>% 
  # filter(minute == 0) %>% 
# filter(time == hm("8:00")) %>% arrange(id) %>% 
  st_drop_geometry() %>% 
  filter(hour >=7 & hour <=10) %>% 
  group_by(fleet) %>% 
  summarise(estimatedDrt_accessibility = sum(estimatedDrt_accessibility*population)/sum(population), car_accessibility = sum(car_accessibility * population)/sum(population), pt_accessibility = sum(pt_accessibility * population)/sum(population), teleportedWalk_accessibility = sum(teleportedWalk_accessibility * population)/sum(population)) %>%
    pivot_longer(ends_with("accessibility"), names_to = "mode", values_to = "accessibility") %>% 
mutate(
  mode = case_when(
    mode == "teleportedWalk_accessibility" ~ "walk",
    mode == "estimatedDrt_accessibility"   ~ "drt",
    mode == "pt_accessibility"             ~ "pt",
    mode == "car_accessibility"            ~ "car",
    TRUE ~ mode
  )
) %>% 
  ggplot() +
  geom_col(aes(fleet,accessibility)) + # fill = hour, position = "dodge"
  facet_grid(~mode) +
  labs(x = "Hour", y = "DRT Accessibility (utils)", title = "Accessibility 7-10AM to Supermarket") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))





```


```{r}
# Prepare max_with_drt and max_without_drt vlaues
equity_distribution_supermarket_800 <- acc_combined %>% 
  filter(time == hm("8:00")) %>% 
  st_drop_geometry() %>% 
  mutate(max_with_drt = pmax(pmax(pt_accessibility,estimatedDrt_accessibility),teleportedWalk_accessibility)) %>%
  mutate(max_without_drt = pmax(pt_accessibility,teleportedWalk_accessibility)) %>% 
  select(population,fleet,`With DRT` = max_with_drt,`No DRT` = max_without_drt) %>%
  uncount(population) %>%
  pivot_longer(ends_with("DRT"), names_to = "scenario",values_to = "value") %>% 
  mutate(scenario = paste0(scenario," (",fleet,")")) %>% 
  filter(scenario == "No DRT (old)" | str_starts(scenario, "With")) %>% 
  mutate(scenario = factor(scenario, levels = c("No DRT (old)","With DRT (old)", "With DRT (small)", "With DRT (medium)", "With DRT (large)")))


lines_and_labels <- equity_distribution_supermarket_800 %>%
  filter(scenario %in% c("No DRT (old)", "With DRT (large)")) %>% 
  group_by(scenario) %>%
  summarise(
    q10    = quantile(value, 0.10, na.rm = TRUE),
    q25    = quantile(value, 0.25, na.rm = TRUE),
    q50 = median(value, na.rm = TRUE),
    q75    = quantile(value, 0.75, na.rm = TRUE),
    q90 = quantile(value, 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  select(-scenario) %>% data.table::transpose(keep.names = "variable") %>% 
  mutate(y1 = as.numeric(str_replace(variable,"q","0."))) %>% 
  mutate(y2 = y1) %>% 
  select(x1=V1, x2=V2, y1,y2) %>% 
  mutate(diff = round(x2-x1,2))
  
# also calculate mean
means <- equity_distribution_supermarket_800 %>% group_by(scenario) %>% summarise(mean = mean(value)) %>% pull(mean)

# plug together plot
plot_ecdf <- ggplot() +
  # ggplot(lines_and_labels) +
  # geom_segment(aes(x = -20, xend = x1, y = y1, yend = y2), color = "gray30", linetype = "dashed", lwd = 0.3) +
  # geom_segment(aes(x = x1, xend = x2, y = y1, yend = y2), color = "gray30") +
  # geom_text(aes(x = x1 - 0.5, y= y2+.05, label = paste0("Δ = ", diff)),hjust=1, inherit.aes = FALSE, size = 4)+
  stat_ecdf(data = equity_distribution_supermarket_800, mapping =  aes(value, col = scenario), linewidth = 1) +
  # scale_color_manual(values = c(`No DRT` = "#1b9e77",`With DRT` = "purple")) +
  labs(x = "Accessibility (utils)", y = "Cumulative Share (≤ x)")+
  theme_minimal(base_size = 12)+ 
  # theme(legend.position = "bottom") +
  # theme( panel.grid.minor.y = element_blank())+
  coord_cartesian(xlim = c(-15, 1))+
  
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2)
  )
  # geom_point(data = data.frame(x = means, y = c(.385,.43)), aes(x,y), col = "red", shape = "diamond", size = 3)+
  # geom_line(data = data.frame(x = means, y = c(.385,.43)), aes(x,y), col = "red") +
  # geom_text(aes(x = means[1] - .5,y = 0.415, label = paste0("Δ = ", round(means[2]-means[1],2))),hjust=1, inherit.aes = FALSE, size = 4, color = "red")

plot_ecdf

```


#XXX


### Kai Analysis


```{r}
comb <- persons_base %>%  filter(subpopulation == "person") %>% st_as_sf(coords = c("home_x","home_y"), crs=25832) %>% st_filter(kelheim_shp) %>% select(person,executed_score) %>% st_join(acc_combined %>% filter(vehicles == 10, time == hm("12:00")),join = st_nearest_feature)  


ggplot(comb) + geom_point(aes(executed_score, car_accessibility), alpha = 0.1, size = 0.1)
ggplot(comb) + geom_point(aes(executed_score, pt_accessibility), alpha = 0.1, size = 0.1)
ggplot(comb) + geom_point(aes(executed_score, teleportedWalk_accessibility), alpha = 0.1, size = 0.1)

cor(comb$executed_score,comb$teleportedWalk_accessibility)


library(purrr)

vars_x <- c("executed_score")


vars_y <- c(
  "car_accessibility","pt_accessibility","teleportedWalk_accessibility"
)

pairs <- expand_grid(var1 = vars_x, var2 = vars_y)

result <- pairs %>%
  mutate(
    test = map2(var1, var2, ~ cor.test(comb[[.x]], comb[[.y]])),
    correlation = map_dbl(test, "estimate"),
    p_value = map_dbl(test, "p.value"),
    stars = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE ~ ""
    ),
    corr_rounded = sprintf("%.3f%s", round(correlation, 3), stars)
  ) %>%
  select(var1, var2, correlation, p_value, corr_rounded)

result
```



### immobile analysis
```{r}
persons_base


# 25,184 adult agents living in Kelheim
persons_base %>% filter(person %in% persons_in_kelheim) %>% filter(age >=18)

# 5,622  immobile adult agents living in kelheim = 22.3%. In the MiD it is 14%
persons_base %>% filter(person %in% persons_in_kelheim) %>% filter(age >=18) %>% left_join(trips_base %>% group_by(person) %>% count()) %>% filter(is.na(n))


# all kelheim adult agents have car availability
persons_base %>% filter(age >=18) %>% group_by(carAvail) %>% count() %>% ungroup() %>% mutate(share = n/sum(n)*100.)




```


# DEPRECATED 
mode stats file
```{r}
mode_stats_small <- read_delim(paste0(output_path_small,"modestats.csv"),delim = ";") %>% 
  pivot_longer(-iteration, names_to = "mode",values_to = "share") %>%  
  filter(mode!="freight") 

mode_stats_small%>% 
  ggplot() + geom_line(aes(iteration,share, col = mode))

mode_stats_large <- read_delim(paste0(output_path_large,"modestats.csv"),delim = ";") %>% 
  pivot_longer(-iteration, names_to = "mode",values_to = "share") %>%  
  filter(mode!="freight") 

mode_stats_large%>% 
  ggplot() + geom_line(aes(iteration,share, col = mode))



mode_stats_small %>% filter(iteration == 300) %>% left_join(mode_stats_large %>% filter(iteration == 300), by = c("iteration","mode")) %>% mutate(diff = share.y - share.x)


# who is using drt - distributed around study area 
# trips_small %>% filter(main_mode == "drt") %>% select(person) %>% left_join(persons_raw) %>% st_as_sf(coords = c("home_x","home_y"), crs = 25832) %>% tm_shape()+ tm_dots()


```


wait time & detour factor 
```{r}
data <- read_delim(paste0(output_path_small,"drt_service_quality_probes.csv.gz"), delim = ";")

data %>% mutate(time = as.factor(time / 3600)) %>% ggplot() + geom_boxplot(aes(time,waitTime))

tmap_mode("view")
data %>% filter(time == 16*3600) %>% group_by(originZone) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor), originX = first(originX), originY = first(originY)) %>% st_as_sf(coords = c("originX","originY"), crs=25832) %>% tm_shape() + tm_dots(fill = "detourFactor", size = 2)

data %>% filter(time == 16*3600) %>% group_by(originZone) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor), originX = first(originX), originY = first(originY)) %>% st_as_sf(coords = c("originX","originY"), crs=25832) %>% tm_shape() + tm_dots(fill = "waitTime", size = 2)


```


```{r}
read_delim("/Users/jakob/git/public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/develop/2026-07-03/output-300it-10veh-1pct-500m/kelheim-v3.1-1pct-kexi-iter_300.drt_service_quality_probes.csv.gz", delim = ";") %>% filter(time == 21600) %>% summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor))

read_delim("/Users/jakob/git/matsim-kelheim/output/output-kelheim-v3.0-1pct-iter_1-plans_kelheim-v3-1-1pct-kexi-iter_300-output_plans/kelheim-v3.1-1pct-kexi-iter_1-plans_kelheim-v3-1-1pct-kexi-iter_300-output_plans.drt_service_quality_probes.csv.gz", delim = ";") %>% view() 

summarise(waitTime = mean(waitTime), detourFactor = mean(detourFactor))
```
