---
title: "postprocess"
output: html_document
date: "2025-06-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(tidyverse)
library(tmap)
library(sf)
library(RColorBrewer)
library(scales)
library(tmaptools)
library(lubridate)
library(hms)
library(RColorBrewer)
library(viridisLite)
library(osmdata)
library(xml2)
library(xtable)
library(patchwork)
```

# INPUTS

## Accessibility File
Remember to set matsim-kelheim as working directory
```{r}
main_time <- "7:30"
# input directory
base_dir <- "../public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/2026-01-12-b/"

acc_dir <- paste0(base_dir,"analysis/accessibility/")
figure_dir <-  paste0(base_dir, "figures-R/")
dir.create(figure_dir)


accessibility_full <- data.frame()
poi_full <- data.frame()
pois <- c("train_station", "logistic", "supermarket")

for(poi in pois){
  
  accessibility_full <- rbind(accessibility_full, read_csv(paste0(acc_dir, poi, "/accessibilities_simwrapper.csv")) %>% mutate (poi = poi))
  poi_full  <- rbind(poi_full,st_read(paste0(acc_dir,poi,"/pois.shp")))
  
}


accessibility_full <-  accessibility_full %>% 
  st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>% 
  mutate(time = as_hms(time))
  
bbox = bb(accessibility_full, ext = 1.05)

```

## Transport Systems

### PT
read stops
Remember to set matsim-kelheim as working directory
```{r}
# read the XML file
doc <- read_xml(paste0(base_dir, "/kexi-seed1-ASC-2.45.output_transitSchedule.xml.gz"))
shp_kelheim <- st_read("input/shp/lk-kelheim/lk-kelheim.shp")

# extract all stopFacility nodes
stops <- xml_find_all(doc, ".//stopFacility")

# turn into a data frame
df <- tibble(
  id = xml_attr(stops, "id"),
  x  = as.numeric(xml_attr(stops, "x")),
  y  = as.numeric(xml_attr(stops, "y")),
  name = xml_attr(stops, "name")
)

tmap_mode("view")
pt_stops <- df %>% st_as_sf(coords = c("x","y"), crs = "EPSG:25832") %>% st_filter(shp_kelheim)
tm_shape(pt_stops) + tm_dots()

```

### CAR
read major roads from osm

```{r}


# query OSM for highways
q <- opq(bbox = bbox %>% st_transform(4326)) %>% 
  add_osm_feature(key = "highway", 
                  value = c("motorway", "trunk", "primary","secondary"))

roads_osm <- osmdata_sf(q)$osm_lines

tm_shape(roads_osm) + tm_lines()
```


### DRT
```{r}

# extract all stopFacility nodes
drt_stops <- st_read("../public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/_data/1_processed/drt_stops/stops_25832.shp")


tm_shape(drt_stops) + tm_dots() + tm_shape(shp_kelheim) + tm_borders()

```

# ACCESSIBILITY CALCULATION

Define Palettes

```{r}
col_title <-  "Color: Accessibility\n[utils]"
# col_vals <-  c("#B2182B", "#D6604D", "#F4A582" ,"#92C5DE", "#4393C3", "#2166AC")

col_vals_div <- brewer.pal(7, "RdYlBu")
col_breaks_div <-  c(-Inf,seq(-7,7,by = 2), Inf)
midpoint_div <- 0

col_vals_pure <- viridis(9)
col_breaks_pure <-  c(-Inf,seq(-14,-2,by = 2), Inf)
midpoint_pure <- -8



alpha_vals <- c(0.0,0.25,0.5,1.0)
alpha_breaks <- c(-Inf,5,50,500,Inf)
```

Function to generate accessibility map for mode, poi, time, and various styling 

```{r}

generate_map <- function(mode1, poi1,time1, opacity, col_vals = NULL, col_breaks = NULL, midpoint = NULL,file_name_append = NULL){
  
  print(paste0("POI:", poi1, " | Time:",time1," | Opacity: ",opacity, " | Mode Combo: ", paste(mode1, collapse = ",")))
  
  # SETUP 
  setup <- accessibility_full %>%
      filter(time == hm(time1)) %>%
    filter(poi == poi1) %>% 
    mutate(population_density = population / (0.5*0.5))
  
  
  if(length(mode1)==2) {
    
    # label <- paste(paste(mode, collapse="-"), "Accessibility")
    label <- "Accessibility\nDiff (util)"
     # print(label)
     setup <- setup %>%  mutate(
       !!label := !!sym(paste0(mode1[[1]], "_accessibility")) -
                               !!sym(paste0(mode1[[2]], "_accessibility")))
     
     col_vals <- if(is.null(col_vals)) col_vals_div else col_vals
     
     col_breaks <- if(is.null(col_breaks)) col_breaks_div else col_breaks
     
     midpoint <- if(is.null(midpoint)) midpoint_div else midpoint
    
  }else{
    # label <- paste(mode, "Accessibility")
    label <- "Accessibility\n(util)"
    setup <- setup %>%  mutate(
       !!label := !!sym(paste0(mode1, "_accessibility"))
    )
    
     col_vals <- if(is.null(col_vals)) col_vals_pure else col_vals
     col_breaks <- if(is.null(col_breaks)) col_breaks_pure else col_breaks
     midpoint <- if(is.null(midpoint)) midpoint_pure else midpoint
    
  }
     
    
    
    # Use this expanded bbox in tm_shape
    
    tmap_mode("plot")
    # https://leaflet-extras.github.io/leaflet-providers/preview/
    
    
    # ADD FILL AND ALPHA TO MAP BASED ON ACCESSIBILITY
    full_map <- tm_basemap("Esri.WorldGrayCanvas") + 
     tm_shape(setup,bbox = bb(setup, ext = 1.05)) +
      tm_symbols(
        fill = label,
        fill.scale = tm_scale_intervals(values = col_vals, breaks = col_breaks, midpoint = midpoint), 
        fill_alpha = ifelse(opacity=="on","population_density",1.0),
        fill_alpha.scale = tm_scale_intervals(values = alpha_vals, breaks = alpha_breaks),
        fill_alpha.legend =  tm_legend(show=FALSE),
        
        size = .14,
        shape = 15,
        
      )
  
    # ADD LEGEND FOR OPACIty
  if(opacity == "on"){
    full_map <- full_map +
        tm_add_legend(
          type = "symbol",
          labels = c("<5", "5-50","50-500",">=500"),
          col = "blue",  # pick a neutral or representative color
          fill_alpha = c(0.0,0.25,0.5, 1.),
          shape = 15,
          title = "Pop. Density\n[person/km²]"
        )
  }
    
    
    # ADD TRANSPORT SYSTEM
    if(length(mode1)==1 & is.null(file_name_append)){
      if(mode1 == "estimatedDrt") {
        
        
        full_map <- full_map +
               # DRT stops as black dots
              tm_shape(drt_stops) +
              tm_symbols(
                shape = 16,
                col = "gray40",
                size = 0.2,
                legend.shape.show = FALSE
              ) 
          
      }
      
      if(mode1 == "pt") {
        
        
        full_map <- full_map +
               # DRT stops as black dots
              tm_shape(pt_stops) +
              tm_symbols(
                shape = 18,       
                col = "blue",
                size = 0.2,
                legend.shape.show = FALSE
              ) 
          
      }
      
      if (mode1 == "car") {
              full_map <- full_map + tm_shape(roads_osm) + tm_lines(col = "gray40", lwd = 1) 
      }
    }
    
    # ADD POI
    full_map <- full_map +
            # POIs as red triangles
          tm_shape(poi_full %>% filter(type == poi1)) +
          tm_symbols(
            shape = 17,          # filled triangle
            col = "red",
            size = 0.2,
            legend.shape.show = FALSE
          )
    
    
    
  
    # ADD LEGEND ENTRY 
    if(length(mode1)==2 | !is.null(file_name_append)) {
            full_map <- full_map +
            tm_add_legend(
            type = "symbol",
            labels = c("Opportunity"),
            col = c( "red"),
            shape = c(17),
            size = 0.5,
            title = "Location"
          ) 
    }else{
    
      full_map <- full_map +
            tm_add_legend(
            type = "symbol",
            labels = c("DRT Stop", "PT Stop","Major Roads","Opportunity"),
            col = c("gray40","blue","gray40", "red"),
            shape = c(16,18,3,17),
            size = 0.5,
            title = "Location"
          ) 
          # tm_scale_bar(position = c("left","bottom")) +
          # tm_compass(position = c("right","top"))
    }
    # add layout
      full_map <- full_map +  
            tm_layout(
              legend.position = tm_pos_out("right","center"),
              inner.margins = c(0,0,0,0),
              outer.margins = 0
              )
      
      
    # file_name <- paste0(figure_dir,ifelse(opacity == "on", "opacity_",""),paste(mode1,collapse="-"),"_",poi1,"_",time1,ifelse(is.null(file_name_append),"",file_name_append), ".pdf")
    # tmap_save(full_map, filename = file_name,   width = 8, height = 6, units = "cm") # ,asp = 0
    


    return(full_map)
    
}


```

## Run Accessibility Calculation

```{r}

# pure
count <- 1
subfigs = c("a","b","c","d","e","f","g","h","i")
for(poi1 in c("train_station","logistic","supermarket")){
  for(mode1 in c("car","pt","estimatedDrt")) {
    full_map <- generate_map(mode1,poi1,main_time,"off")
    file_name <- paste0(figure_dir,"fig3",subfigs[count],"_",mode1,"_",poi1,".pdf")
    tmap_save(full_map, filename = file_name,   width = 8, height = 6, units = "cm") # ,asp = 0
    count <- count + 1
    
  }
}


# temporal (difference plots DRT - PT)
count <- 1
for(poi in c("train_station")){
  for(time1 in times <- c(main_time,"8:30","24:00")){
    full_map <- generate_map(c("estimatedDrt","pt"), poi, time1, "off")
    file_name <- paste0(figure_dir,"fig4",subfigs[count],"_",paste(mode1,collapse="-"),"_",poi1,"_",time1, ".pdf")
    tmap_save(full_map, filename = file_name,   width = 8, height = 6, units = "cm") # ,asp = 0
    count <- count + 1
  }
}


```

# EQUITY CALCULATION

Prep

```{r}

# We scale up population from 25% scenario to 100% by multiplying by 4
accessibility_processed <- accessibility_full %>% 
  filter(time != hm("8:30") & time !=hm("7:00")) %>% 
  mutate(population = population * 4) %>% 
  mutate(population_density = population / (0.5*0.5))



```


## ECDF Plot
```{r}

# Prepare max_with_drt and max_without_drt vlaues
equity_distribution_supermarket_730 <- accessibility_processed %>% 
  filter(poi == "supermarket") %>% 
  filter(time == hm("7:30")) %>% 
  st_drop_geometry() %>% 
  mutate(max_with_drt = pmax(pmax(pt_accessibility,estimatedDrt_accessibility),teleportedWalk_accessibility)) %>%
  mutate(max_without_drt = pmax(pt_accessibility,teleportedWalk_accessibility)) %>% 
  select(population,`With DRT` = max_with_drt,`No DRT` = max_without_drt) %>% uncount(population) %>% 
  pivot_longer(everything(), names_to = "scenario",values_to = "value")


# some values used for paper
equity_distribution_supermarket_730  %>% mutate(below6 = value < -6, below3 = value < -3) %>% group_by(scenario) %>% summarise(below6 = sum(below6)/n(),below3 = sum(below3)/n())
  

# calculate the percentiles, and create labels used for ecdf
lines_and_labels <- equity_distribution_supermarket_730 %>%
  group_by(scenario) %>%
  summarise(
    q10    = quantile(value, 0.10, na.rm = TRUE),
    q25    = quantile(value, 0.25, na.rm = TRUE),
    q50 = median(value, na.rm = TRUE),
    q75    = quantile(value, 0.75, na.rm = TRUE),
    q90 = quantile(value, 0.90, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  select(-scenario) %>% data.table::transpose(keep.names = "variable") %>% 
  mutate(y1 = as.numeric(str_replace(variable,"q","0."))) %>% 
  mutate(y2 = y1) %>% 
  select(x1=V1, x2=V2, y1,y2) %>% 
  mutate(diff = round(x2-x1,2))
  
# also calculate mean
means <- equity_distribution_supermarket_730 %>% group_by(scenario) %>% summarise(mean = mean(value)) %>% pull(mean)

# plug together plot
plot_ecdf <- ggplot(lines_and_labels) +
  geom_segment(aes(x = -20, xend = x1, y = y1, yend = y2), color = "gray30", linetype = "dashed", lwd = 0.3) +
  geom_segment(aes(x = x1, xend = x2, y = y1, yend = y2), color = "gray30") +
  geom_text(aes(x = x1 - 0.5, y= y2+.05, label = paste0("Δ = ", diff)),hjust=1, inherit.aes = FALSE, size = 4)+
  stat_ecdf(data = equity_distribution_supermarket_730, mapping =  aes(value, col = scenario), linewidth = 1) +
  scale_color_manual(values = c(`No DRT` = "#1b9e77",`With DRT` = "purple")) +
  labs(x = "Accessibility (utils)", y = "Cumulative Share (≤ x)")+
  theme_minimal(base_size = 14)+ 
  theme(legend.position = "bottom") +
  # theme( panel.grid.minor.y = element_blank())+
  coord_cartesian(xlim = c(-15, 1))+
  
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2)
  )+ geom_point(data = data.frame(x = means, y = c(.385,.43)), aes(x,y), col = "red", shape = "diamond", size = 3)+
  geom_line(data = data.frame(x = means, y = c(.385,.43)), aes(x,y), col = "red") +
  geom_text(aes(x = means[1] - .5,y = 0.415, label = paste0("Δ = ", round(means[2]-means[1],2))),hjust=1, inherit.aes = FALSE, size = 4, color = "red")

plot_ecdf

# save plot
ggsave(plot_ecdf, file = paste0(figure_dir,"fig5_equity_ecdf.pdf"),width = 10, height = 10, units = "cm", dpi = 300, device = cairo_pdf)

```

## Smoothed Line Chart
```{r}
# prep
scatter_prep <- accessibility_processed %>% 
  st_drop_geometry() %>% 
  filter(poi == "supermarket") %>%
  select(ends_with("accessibility"),population_density,time,poi) %>% 
  pivot_longer(-c(population_density,time,poi),names_to = "mode", values_to = "accessibility") %>%
  mutate(mode = str_replace(mode,"_accessibility","")) %>% 
  filter(time == hm(main_time)) %>% 
  # filter(mode == "pt" | time == hm(main_time)) %>% 
  mutate(mode = case_when(
    mode == "teleportedWalk" ~ "Walk",
    mode == "estimatedDrt" ~ "DRT",
    mode == "pt" ~ "PT",
    mode == "car" ~ "Car",
    .default = mode))


# create plot
  scatter_supermarket <- ggplot(scatter_prep, aes(population_density,accessibility,col = mode)) +
  # geom_point(alpha = 0.1,size = .3)+
  geom_smooth(method = "loess")+
  scale_x_log10()+
  coord_cartesian(xlim = c(NA, 5000))+
  labs(x="Population Density (p/km2)", y = "Accessibility (utils)") +
      # geom_hline(yintercept = -6.0,
      #        linetype = "dashed",
      #        color = "red",
      #        size = 1) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "bottom")
scatter_supermarket

scatter_prep %>% filter(mode == "walk")  %>% filter(population_density > 3000) %>% nrow()

  # coord_cartesian(ylim = c(-50, 0))
  
  ggsave(filename = paste0(figure_dir,"fig6_equity_scatter.pdf"), plot = scatter_supermarket, width = 10, height = 10, units = "cm", dpi = 300)

  
  # Find some numbers for the paper
  scatter_prep %>% filter(population_density > 0 & population_density <= 50) %>% group_by(mode) %>% summarise(mean = sum(accessibility * population_density)/sum(population_density))
  
  scatter_prep %>% filter(population_density >= 1000) %>% group_by(mode) %>% summarise(mean = sum(accessibility* population_density)/sum(population_density))


```


# Map of areas above threshold

```{r}

for(threshold in c(-6,-3)){
  supermarket_threshold <- accessibility_processed %>%
    filter(time == hm("7:30"), poi == "supermarket") %>% 
    filter(population_density > 5) %>% 
    select(teleportedWalk_accessibility,pt_accessibility,estimatedDrt_accessibility,population,population_density) %>% 
    mutate(walk = teleportedWalk_accessibility >=threshold,
           pt = pt_accessibility >= threshold,
           drt = estimatedDrt_accessibility >= threshold, 
           none = !walk & !pt & !drt)
  
  supermarket_threshold %>% st_drop_geometry() %>% 
    pivot_longer(c("walk","pt","drt","none"),names_to = "scenario") %>% 
    filter(value) %>% 
    group_by(scenario) %>% 
    summarise(population = sum(population)) %>% mutate(scenario = factor(scenario, levels = c("walk","pt","drt","none"))) %>% arrange(scenario) %>% 
    mutate(diff = population - lag(population))
  
  
  subsume_check <- supermarket_threshold %>% 
    mutate(pt_subsumes_walk = !walk | (walk & pt)) %>% 
    mutate(drt_subsumes_walk = !walk | (walk & drt)) %>% 
    mutate(drt_subsumes_pt = !pt | (pt & drt))
  
  
    subsume_check %>% summarise(pt_subsumes_walk =  n()- sum(pt_subsumes_walk), drt_subsumes_walk = n() - sum(drt_subsumes_walk),drt_subsumes_pt =  n() -sum(drt_subsumes_pt))
    
  
  
  equity_map <- tm_basemap("Esri.WorldGrayCanvas") +
    tm_shape(bbox = bb(supermarket_threshold, ext = 1.05)) +
    tm_shape(supermarket_threshold %>% filter(drt)) +
    tm_dots(fill = "purple") +
    tm_shape(supermarket_threshold %>% filter(pt)) +
    tm_dots(fill = "#1b9e77") +
    tm_shape(supermarket_threshold %>% filter(walk)) +
    tm_dots(fill = "yellow") +
    tm_shape(supermarket_threshold %>% filter(none)) +
    tm_dots(fill = "red") 
  
    file_name <- paste0(figure_dir,"fig7",ifelse(threshold == -6, "a","b"),"_equity_map_",threshold,".pdf")
    tmap_save(equity_map, filename = file_name, units = "cm") # ,asp = 0
    
}


legend_only <- tm_add_legend(
  type = "symbol",
  labels = c("Walk","PT", "DRT", "No Access"),
  fill = c("yellow","#1b9e77","purple", "red"),
  lwd = 0,
  size = .4,
  title = "Mode"
) +
  tm_layout(
    legend.only = TRUE,
    frame = FALSE,
    asp = 0,    
    outer.margins = 0,
    inner.margins = 0,
    legend.text.size = 0.9,
    legend.title.size = 1
  )

 file_name <- paste0(figure_dir,"fig7c_equity_map_legend.pdf")
tmap_save(
  legend_only,
  filename = file_name,
  width = 2.3,
  height = 2,
  units = "cm"
)



```


# SENSITIVITY ANALYSIS

```{r}


# prepare reference data
ref_data <- accessibility_full %>% 
  filter(time == hm(main_time)) %>% 
  filter(poi == "train_station") %>% 
  summarize(pt = sum(pt_accessibility * population)  / sum(population), 
            car = sum(car_accessibility*population)  / sum(population), 
            drt = sum(estimatedDrt_accessibility*population)  / sum(population)) %>%
  st_drop_geometry() %>% 
 
  pivot_longer(cols = -c(),names_to = "mode",values_to = "accessibility") %>% 
  mutate(scenario = mode) %>%
  select(scenario, accessibility) %>%
  mutate(scenario = factor(scenario,levels = c("car","drt","pt"),labels = c("Car","DRT","PT")))
  
# grab a column with the populatin, scale up from 25% scenario to 100% scenario
pop_col <- accessibility_full %>% 
  filter(time == hm(main_time)) %>% 
  filter(poi == "train_station") %>% 
  st_drop_geometry() %>% 
  transmute(population = population * 4)


scenarios <- c("a-1","a-2","a-3","b-1","b-2","b-3","c-1","c-2","c-3")

sensitivity_data_raw <- map_dfr(scenarios, function(scenario) {
  path <- paste0(
    "../public-svn/matsim/scenarios/countries/de/kelheim/drtAccessibility/2026-01-09-calibration/",
    scenario,
    "/analysis/accessibility/train_station/accessibilities.csv"
  )

  read_csv(path) %>%
    filter(time == hm(main_time)) %>%
    cbind(pop_col) %>%
    summarize(
      scenario = scenario,
      accessibility = sum(estimatedDrt_accessibility * population) / sum(population)
    )
})

sensitivity_data <- sensitivity_data_raw %>% 
  separate(scenario, into = c("asc","los"), sep = "-") %>% 
  mutate(los = factor(los, levels=c("3","2","1"), labels = c("poor","base","car"))) %>% 
  mutate(asc = factor(asc, levels = c("a","b","c"), labels = c("-2.5","base","2.5"))) 


ref_plot <- ggplot(ref_data) + 
  geom_col(aes(scenario,accessibility), fill = "gray70")+
  labs(x= "Ref Mode", y = "Accessibility")+
  theme_minimal(base_size = 14)+
  theme(
    axis.title.x = element_text(),
    axis.title.y = element_text(),
    axis.text.y  = element_text(),
    axis.text.x  = element_text(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(5, 2, 5, 5)  # remove right margin
  ) +
  coord_cartesian(ylim = c(0,-16))+
  scale_x_discrete(position = "top")

sensitivity_plot <- ggplot(sensitivity_data) + 
  geom_col(aes(asc,accessibility,fill = los, group = los),position = "dodge") +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(5, 5, 5, 2)
  ) +
  coord_cartesian(ylim = c(0,-16)) +
  scale_x_discrete(position = "top")+
  labs(x = "ASC for DRT", fill = "LOS of DRT")



final <- (ref_plot | sensitivity_plot) +
  plot_layout(widths = c(1, 3))

final

ggsave(plot = final,filename = paste0(figure_dir,"fig10_appendix_sensitivity_bar.pdf"), width = 18, height = 10, units = "cm", dpi = 300)


  
```

